Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-10-06 demystify.tex
updates

    % Title
    \node[align=center] at ($(current page.center)+(0,-7)$)
    {
    {\fontsize{24}{24} \selectfont {{Demystifying the Computer Science}}} \\[0.15in]
    {\fontsize{24}{24} \selectfont {{PhD Admission in the US}}} \\[0.2in]    
    {\fontsize{18}{18} \selectfont {{A Handbook for International Students}}} \\[1.in]

@@

% though \emph{very} top schools could be very selective, e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU
So I want to share some details about the admission process and advice for those who are interested in applying for a \textbf{PhD in Computer Science in the US}.
Originally, this document was intended for international students, but I have expanded it to include information that might also be useful for \emph{US domestic students}.
Moreover, while this is primarily intended for students interested in CS, it might be relevant to students from various STEM (Science, Technologies, Engineering, and Mathematics) disciplines.
Furthermore, although many examples are specifics for schools that I and other contributors of this document know about, the information should be generalizable to other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines. Currently, 146 (out of 4000) US universities are classified as R1.} institutions in the US.

@@

I wish you the best of luck. Happy school hunting!

\begin{mybox}
This document will be updated regularly to reflect the latest information and updates in the admission process. Its latest version is available at

@@

First, applying to a good US university \emph{should not} be any harder than at schools in other countries. In fact, it might even be more flexible since CS PhD in the US \emph{do not} require having an MS or a research topic/proposal/adviser in advance. If you believe you have a chance in other countries, e.g., South Korea, Singapore, Germany, UK, Japan, and Australia, then you will surely have a chance in the US as well. \autoref{sec:non-us-differences} compares CS PhD study in the US to other countries.

Second, PhD students in CS \emph{do not} need to worry about funding, especially at good R1
universities in the US. If you are admitted, you will almost certainly
receive \emph{full funding} (\autoref{sec:funding}) to support your study, including tuition,
health insurance, and stipend (i.e., you get paid for your study). Moreover, depending on the university,
you may even receive additional benefits such as summer pay, laptops, and traveling to conferences and workshops.

@@

Students from top schools with strong research cultures and experiences might already know most of the information given here.
My goal is thus to level the playing field by providing information that is not readily available to less privileged students. 
I hope to encourage more students with such backgrounds to apply to CS PhD programs in the US.

@@

Typically, it takes 5--7 years for CS PhD in the US.  This can be longer than CS PhDs in other countries, which might require MS first (recall from \autoref{sec:msrequirement} that CS PhD programs in the US do not require MS and you can get MS along the way to PhD). Within these 5--7 years, CS PhD students often take a ``leave of absence'' for 1--2 semesters to do internships at companies and research labs.

@@

% \begin{table}[h!]
% \centering
% \begin{tabular}{l|l}
%     \toprule
% \textbf{Month} & \textbf{Task} \\
% \midrule
% January-March & Research programs and prepare for tests \\
% April-June & Take standardized tests (GRE, TOEFL, etc.) \\
% July-September & Prepare application materials (SOP, LORs, etc.) \\
% October-December & Submit applications \\
% January-March & Interviews and follow-ups \\
% April & Admission decisions \\
% \bottomrule
% \end{tabular}
% \caption{Application Timeline}
% \end{table}
% \section{Why the US and not other countries?}
% TODO
\subsubsection{Why CS depts do not waive the application fee?}

@@

>>>>

    % Title
    \node[align=center] at ($(current page.center)+(0,-7)$)
    {
    {\fontsize{24}{24} \selectfont {{Demystifying the Computer Science}}} \\[0.15in]
    {\fontsize{24}{24} \selectfont {{PhD Admission in the US}}} \\[1in]    
    %{\fontsize{18}{18} \selectfont {{A Handbook for International Students}}} \\[1.in]

@@

% though \emph{very} top schools could be very selective, e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU
So I want to share some details about the admission process and advice for those who are interested in applying for a \textbf{PhD in Computer Science in the US}.
Originally, this document was intended for \emph{international students}, but I have expanded it to include information that might also be useful for \emph{US domestic students}.
Moreover, while this primarily aims at students interested in CS, it might be relevant to students from various STEM (Science, Technologies, Engineering, and Mathematics) disciplines.
Furthermore, although many examples are specifics for schools that I and other contributors of this document know about, the information should be generalizable to other good R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines. Currently, 146 (out of 4000) US universities are classified as R1.} institutions in the US.

@@

I wish you the best of luck.

\begin{mybox}
This document is open source and updated regularly. Its latest version is available at

@@

First, applying to a good US university \emph{should not} be any harder than at schools in other countries. In fact, it might even be more flexible since CS PhD in the US \emph{do not} require having an MS or a research topic, proposal, or adviser in advance (\autoref{sec:non-us-differences} compares CS PhD study in the US to other countries). If you believe you have a chance in other countries, e.g., South Korea, Singapore, Germany, UK, Japan, and Australia, then you will surely have a chance in the US as well.

Second, PhD students in CS \emph{do not} need to worry about funding, especially at good R1
universities in the US. If you are admitted, you will almost certainly
receive \emph{full funding} (\autoref{sec:funding}) to support your study. This includes tuition,
health insurance, and stipend (i.e., you get paid for your study). Moreover, depending on the university,
you may receive additional benefits such as summer pay, laptops, and traveling to conferences and workshops.

@@

Students from top schools with strong research cultures and experience might already know most of the information given here.
My goal is thus to level the playing field by providing information that is not readily available to less privileged students. 
I hope to encourage more students with such backgrounds to apply to CS PhD programs in the US.

@@

Typically, it takes 5--7 years for CS PhD in the US.  This can be longer than CS PhDs in other countries (\autoref{sec:msrequirement}), which might require MS first. Within these 5--7 years, CS PhD students often take a ``leave of absence'' for 1--2 semesters to do internships at companies and research labs.

@@

% \begin{table}[h!]
% \centering
% \begin{tabular}{l|l}
%     \toprule
% \textbf{Month} & \textbf{Task} \\
% \midrule
% January-March & Research programs and prepare for tests \\
% April-June & Take standardized tests (GRE, TOEFL, etc.) \\
% July-September & Prepare application materials (SOP, LORs, etc.) \\
% October-December & Submit applications \\
% January-March & Interviews and follow-ups \\
% April & Admission decisions \\
% \bottomrule
% \end{tabular}
% \caption{Application Timeline}
% \end{table}
% \section{Why the US and not other countries?}
% TODO

@@

\subsection{Why CS depts do not waive the application fee?}
